---
title: "Title"
author: "Author Names"
date: "Date"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, message=FALSE, warning=FALSE, fig.height=4, fig.width=10)
```


```{r}
# load packages - add any you would like
library(tidyverse)
library(lme4)
library(lmerTest)
library(gridExtra)
library(knitr)
```

## Abstract 

A 150 word abstract.


\newpage

### Background and Significance   

Type background and significance section here. 

### Data  

Type data section here. Include captions with tables and graphs. Examples:  

```{r}
music <- read.csv("https://raw.githubusercontent.com/proback/BeyondMLR/master/data/musicdata.csv")
T1 <- head(music %>% select(id, perform_type, memory, na, age, gender, instrument, years_study, mpqpem, mpqnem, mpqab ),10)       
kable(T1, caption="First 10 rows of the dataset")
```


```{r, fig.cap="Negative affect (a) for all 497 performances and (b) average for each musician "}
# Create Level2 data set by picking off one observation 
# per subject, which would be easier if every subject 
# had a diary entry labeled '1' - should be 37 rows 
# and 6 columns (one per L2 variable)
music.lev2 <-  music %>%
  group_by(id) %>%
  filter(row_number() == 1) %>%
  select(id, gender:mpqnem)
# Add average across all performances for each subject 
# for EDA plots
meanbysubj <- music %>% group_by(id) %>%
  summarise(meanbysubj = mean(na, na.rm = TRUE))
music.lev2 <- music.lev2 %>%
  left_join(meanbysubj, by = "id")

p1 <- ggplot(data=music,aes(x=na)) + 
  geom_histogram(binwidth = 2, fill = "white",color = "black") + 
  xlab("Negative Affect") + ylab("Frequency") + labs(title="(a) Individual Performances") 
p2 <- ggplot(data=music.lev2,aes(x=meanbysubj)) + 
  geom_histogram(binwidth = 2, fill = "white",color = "black") + 
  xlab("Negative Affect") + ylab("Frequency") + labs(title="(b) Musician Averages") 
grid.arrange(p1, p2, ncol=2)
```



### Methods  

Type methods section here. Write your final model in equation form, like this:


\begin{align*}
Y_{ij} & = [\alpha_{0}+\alpha_{1}\textrm{Orch}_{i}+\beta_{0}\textrm{LargeEns}_{ij}+\beta_{1}\textrm{Orch}_{i}\textrm{LargeEns}_{ij}] \\
    & + [u_{i}+v_{i}\textrm{LargeEns}_{ij}+\epsilon_{ij}]
\end{align*}

where,

$$
\left[ \begin{array}{c}
            u_{i} \\ v_{i}
          \end{array}  \right] \sim N \left( \left[
          \begin{array}{c}
            0 \\ 0
          \end{array} \right], \left[
          \begin{array}{cc}
            \sigma_{u}^{2} & \rho_{uv}\sigma_{u}\sigma_v \\
            \rho_{uv}\sigma_{u}\sigma_v & \sigma_{v}^{2}
          \end{array} \right] \right) 
$$

and $\epsilon_{ij}\sim N(0,\sigma^2)$.   



### Results  

Write your results section here. Include the fixed and random effects tables, as below. 

```{r}
music <- music %>% mutate(orch = as.numeric(instrument=="orchestral instrument"),
                          large = as.numeric(perform_type=="Large Ensemble"))
M <- lmer(na ~ orch + large + orch:large +
  (large|id), data = music)
```

```{r}
kable(as.data.frame(VarCorr(M)), caption="Variance and Covariance Estimates for Random Effects")
```


```{r}
kable(summary(M)$coefficients, caption="Estimates of Fixed Effects")
```


### Discussion and Conclusions   


Type discussion section. 

\newpage 

### References 

References in APA format. For example:

Sadler, M. E., & Miller, C. J. (2010). Performance anxiety: A longitudinal study of the roles of personality and experience in musicians. Social Psychological and Personality Science, 1(3), 280-287.

Lin, M. C. (2019). An Investigation Of Music Performance Anxiety In Taiwanese Pianists, Vocalists, String And Wind Instrumentalists At The College Level.

Stoeber, J., & Eismann, U. (2007). Perfectionism in young musicians: Relations with motivation, effort, achievement, and distress. Personality and Individual Differences, 43(8), 2182-2192.

Roback, P., & Legler, J. (2021). Beyond Multiple Linear Regression: Applied Generalized Linear Models And Multilevel Models in R. CRC Press.



\newpage

## Appendix (optional)

